---
title: "Yeast SNPro Phylo"
author: "K Harrison"
date: "2022-07-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Load libraries
```{r}
===
library(phyloseq)
library(devtools)
library(dplyr)
library(vegan)
library(ggplot2)
library(ape)
library(ggpubr)
library(readr)
library(devtools)
library(ggfortify)
library(ggforce)
library(ellipse)
library(data.table)
#library(decontam)
library(microViz)
install.packages(c("devtools", "RcppEigen", "RcppParallel", "Rtsne", "ggforce", "units"))
install.packages("remotes")
remotes::install_github("vmikk/metagMisc")
library(metagMisc)
```

```


```{r}
setwd("~/Documents/Data/Probiotic SN Study/Packaged product analysis/R data")
#Yeast
asv_matITS<- read_csv("ITS ASV SNPro Pro.csv")
#ASV counts
tax_matITS<- read_csv("ITS TAXA SNPro.csv") 
#Taxa comparison
samples_dfITS <- read_csv("ITS SAMP SNPro Pro.csv")
###set demo info to factors so that we can merge downstream
samples_dfITS$Type <-as.factor(samples_dfITS$Type)
samples_dfITS$Claim <-as.factor(samples_dfITS$Claim)
samples_dfITS$Project <-as.factor(samples_dfITS$Project)

#------#define the row names from the otu column- bacteria
asv_matITS <- asv_matITS %>%
  tibble::column_to_rownames("ASV") 
asv_matITS <- as.matrix(asv_matITS) #idem matrix

tax_matITS <- tax_matITS %>% 
  tibble::column_to_rownames("ASV")
tax_matITS <- as.matrix(tax_matITS)

samples_dfITS <- samples_dfITS %>% 
  tibble::column_to_rownames("ID")
```

```{r}
#Phyloseq object 
ASV.ITS = otu_table(asv_matITS, taxa_are_rows = TRUE)
TAX.ITS = tax_table(tax_matITS)
samples.ITS = sample_data(samples_dfITS)
df.fungi <- phyloseq(ASV.ITS, TAX.ITS, samples.ITS)
df.fungi
```


```{r}
df.fungi <- subset_taxa(df.fungi, Genus != "Fungi_phylum_unknown")
df.fungi<- subset_samples(df.fungi, NMR.Data == "Yes")
# Determine which genera are present
#tax_table(df.fungi)[, "Species"]

#generate random tree and merge into phyloseq object for ordinate calculations
random_tree = rtree(ntaxa(df.fungi), rooted=TRUE, tip.label=taxa_names(df.fungi))
#plot(random_tree)
df.fungi = merge_phyloseq(df.fungi, random_tree)

df.fungi.prod <- subset_samples(df.fungi, Project == "Package Product")
#tax_table(df.bacteria.prod)[, "Genus"]
#create a dataset for the packaged product analysis
df.fungi.fil = filter_taxa(df.fungi, function(x) sum(x > 5) > (0.25*length(x)), TRUE) ##Remove taxa not seen more than 5 times in at least 25% of the samples. This protects against an OTU with small mean & trivially large C.V.
#df.bacteria <- tax_glom(df.bacteria.fil, taxrank = "Genus") #merge genus-level
#tax_table(df.bacteria.fil)[, "Genus"]
RA.fungi<- transform_sample_counts(df.fungi.fil, function(x) x / sum(x) )
tax_table(RA.fungi)
#merge species into genera
RA.fungi.genus = tax_glom(RA.fungi, "Genus")

yeast.RA.matrix <- phyloseq_to_df(RA.fungi, addtax = T, addtot = T, addmaxrank = F,
  sorting = "abundance")
  
yeast.RA.matrix.g <- phyloseq_to_df(RA.fungi.genus, addtax = T, addtot = T, addmaxrank = F,
  sorting = "abundance")

write.csv(yeast.RA.matrix.g, "RA.ITS.matrix.g.csv")
```



```{r}
p1 <- plot_bar(RA.fungi, fill = "Species", title = "") + facet_wrap(~Type, scales= "free_x", nrow=1) + geom_point() + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
p1 = p1 + geom_bar(aes(color=Species, fill=Species), stat="identity", position="stack")

p2 <- plot_bar(RA.fungi, fill = "Genus", title = "") + facet_wrap(~Type, scales= "free_x", nrow=1) +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
p2 = p2 + geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") #+ theme(legend.position = "none") 

p3 <- plot_bar(RA.fungi, fill = "Species", title = "") + facet_wrap(~Claim, scales= "free_x", nrow=1) +
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
p3 = p3 + geom_bar(aes(color=Species, fill=Species), stat="identity", position="stack") #+ theme(legend.position = "none") 

p4 <- plot_bar(RA.fungi, fill = "Genus", title = "") + facet_wrap(~Claim, scales= "free_x", nrow=1) +
  theme(axis.ticks.x = element_blank(),
        axis.text.x = element_blank())
p4 = p4 + geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") #+ theme(legend.position = "none") 

ggarrange(p2,p4, nrow = 1,ncol=2,common.legend = TRUE, legend="right")
```
ggarrange(p1,p2, nrow = 1,ncol=2, legend="right")

Alpha diversity-Plots
```{r}
sample_data(df.fungi.fil)$simpson.physeq <- estimate_richness(df.fungi, measures="Simpson")
p <- plot_richness(df.fungi.fil, "Claim", measures=c("Chao1", "Shannon"), color="Claim") + theme_bw()
p + geom_boxplot(data=p$data, aes(x=Claim, y=value, color=NULL), alpha=0.1)
```
p <- plot_richness(df.fungi.fil, "Type", measures=c("Chao1", "Shannon"), color="Type") + theme_bw()
p
p + geom_boxplot(data=p$data, aes(x=Type, y=value, fill=Type), alpha=0.2)

```{r}
#Measure Alpha diversity measure- Shannon
df.fungi.shannon <- estimate_richness(df.fungi.fil, measures="Shannon")

#Measure Alpha diversity measure- Chao1
df.fungi.chao1 <- estimate_richness(df.fungi.fil, measures="Chao1")

#Plot alpha diversity by Type, Claim, PStrain.Listed
p.alpha.claim <- plot_richness(df.fungi.fil, "Claim", measures=c("Chao1", "Shannon"), color = "Claim") 
p.alpha.type <- plot_richness(df.fungi.fil, "Type", measures=c("Chao1", "Shannon"), color = "Type") 
 
p.alpha.claim + geom_boxplot(alpha=0.3)
p.alpha.type + geom_boxplot(alpha=0.3)
#Pairwise Wilcoxon rank-sum test for Beta Diversity
pairwise.wilcox.test(df.fungi.chao1$Chao1,sample_data(df.fungi.fil)$Type)
#              Hard  Soft 
#Soft          0.039 -    
#Soft-Aberrant 0.230 0.562
pairwise.wilcox.test(df.fungi.chao1$Chao1,sample_data(df.fungi.fil)$Claim)
#         live   none  
#none      0.0096 -     
#probiotic 0.0982 0.0920

pairwise.wilcox.test(df.fungi.shannon$Shannon,sample_data(df.fungi.fil)$Type)
#              Hard Soft
#Soft          0.18 -   
#Soft-Aberrant 0.67 0.10
pairwise.wilcox.test(df.fungi.shannon$Shannon,sample_data(df.fungi.fil)$Claim)
#          live  none 
#none      0.035 -    
#probiotic 0.151 0.035
```


```{r}
df.fungi.fil <- tax_fix(df.fungi.fil)
df.fungi.fil <- phyloseq_validate(df.fungi.fil, remove_undetected = TRUE)
df.fungi.fil %>% 
  tax_transform(trans = "clr", rank = "Genus")


df.fungi.fil %>% 
  tax_transform("clr", rank = "Genus") %>% 
  # when no distance matrix or constraints are supplied, PCA is the default/auto ordination method
  ord_calc() %>% 
  ord_plot_iris(tax_level = "Genus", anno_colour = "Type")


#df.fungi.fil %>% 
#  tax_transform("clr", rank = "Genus") %>% 
#  # when no distance matrix or constraints are supplied, PCA is the default/auto ordination method
#  ord_calc() %>% 
#  ord_plot_iris(tax_level = "Genus", ord_plot = "above", anno_colour = "Type")



#scree
DistUF = distance(df.fungi.fil, method = "bray")
ordUF = ordinate(df.fungi.fil, method = "PCoA", distance = DistUF)
plot_scree(ordUF, "Scree Plot: Weighted UniFrac MDS") + theme_bw()

adonis2(DistUF ~ sample_data(df.fungi.fil)$Type) # 0.052*
adonis2(DistUF ~ sample_data(df.fungi.fil)$Claim) #0.013 *

plot_ordination(df.fungi.fil, ordinate(df.fungi.fil, method='NMDS', dist = 'bray'), type='split',
                color='Type')
ord <- ordinate(df.fungi.fil, method='NMDS')
p <- plot_ordination(df.fungi.fil, ord, type='sample', 
                color='Type')
#t-distribution and normal distribution         
p + 
  stat_ellipse(type = "norm", linetype = 2) +
  stat_ellipse(type = "t") + 
  theme_bw()


```

```{r}
df.fungi <- phyloseq_to_df(RA.fungi, addtax = T, addtot = F, addmaxrank = F,
  sorting = "abundance")
write.csv(df.fungi,"~/Documents/Data/Probiotic SN Study/Packaged product analysis/R data/dffungi.csv", row.names = TRUE)

```


By species
```{r}
RA.ANOVA<- read_csv("RA.ITS.matrix.csv") #ASV counts

kruskal.test(`Brettanomyces bruxellensis`~ Type, data = RA.ANOVA) 
#p-value = 0.3714

kruskal.test(`Saccharomyces_species_unknown`~ Type, data = RA.ANOVA) 
#p-value = 0.01786
dunnTest(`Saccharomyces_species_unknown`~ Type, data = RA.ANOVA, method="bh") 
#Hard - Soft-Aberrant p = 0.01608394

kruskal.test(`Hanseniaspora valbyensis`~ Type, data = RA.ANOVA) 
#p-value = 0.346

kruskal.test(`Lachancea fermentati`~ Type, data = RA.ANOVA) 
#p-value = 0.8357

kruskal.test(`Saccharomycetaceae_genus_unknown`~ Type, data = RA.ANOVA) 
#p-value = 0.01786

kruskal.test(`Brettanomyces anomala`~ Type, data = RA.ANOVA) 
#p-value = 0.03064
dunnTest(`Brettanomyces anomala`~ Type, data = RA.ANOVA, method="bh") 
#Hard - Soft p-value = 0.02860248

kruskal.test(`Brettanomyces anomala`~ Type, data = RA.ANOVA) 
#p-value = 0.03064
dunnTest(`Brettanomyces anomala`~ Type, data = RA.ANOVA, method="bh") 
#Hard - Soft p-value = 0.02860248

kruskal.test(`Brettanomyces anomala`~ Type, data = RA.ANOVA) 
#p-value = 0.03064
dunnTest(`Brettanomyces anomala`~ Type, data = RA.ANOVA, method="bh") 
#Hard - Soft p-value = 0.02860248

kruskal.test(`Issatchenkia orientalis`~ Type, data = RA.ANOVA) 
#p-value = 0.8253

kruskal.test(`Kregervanrija fluxuum`~ Type, data = RA.ANOVA) 
#p-value = 0.3696
```

By genera
```{r}
RA.ANOVA<- read_csv("RA.ITS.matrix.g.csv") #ASV counts

kruskal.test(`Brettanomyces`~ Type, data = RA.ANOVA) 
#p-value = 0.1367

kruskal.test(`Saccharomyces`~ Type, data = RA.ANOVA) 
#p-value = 0.04622
dunnTest(`Saccharomyces`~ Type, data = RA.ANOVA, method="bh") 
#Hard - Soft-Aberrant, p-vale = 0.04544316

kruskal.test(`Hanseniaspora`~ Type, data = RA.ANOVA) 
#p-value = 0.346
kruskal.test(`Lachancea`~ Type, data = RA.ANOVA) 
#p-value = 0.8357
kruskal.test(`Hanseniaspora`~ Type, data = RA.ANOVA) 
#p-value = 0.346
kruskal.test(`Saccharomycetaceae_genus_unknown`~ Type, data = RA.ANOVA) 
#p-value = 0.1307
kruskal.test(`Issatchenkia`~ Type, data = RA.ANOVA) 
#p-value = 0.8253
kruskal.test(`Pichiaceae_genus_unknown`~ Type, data = RA.ANOVA) 
#p-value = 0.4677
kruskal.test(`Starmerella`~ Type, data = RA.ANOVA) 
#p-value = 0.3232
kruskal.test(`Cryptococcus`~ Type, data = RA.ANOVA) 
#p-value = 0.6053
```

Looking at replicate samples 
```{r}
#Yeast
asv_matITS.R<- read_csv("ITS ASV SNPro.csv") #ASV counts
tax_matITS.R<- read_csv("ITS TAXA SNPro.csv") #Taxa comparison
samples_dfITS.R <- read_csv("ITS SAMP SNPro.csv") #Sample comparison
###set demo info to factors so that we can merge downstream
samples_dfITS.R$Replicated <-as.factor(samples_dfITS.R$Replicated)
samples_dfITS.R$Replicate <-as.factor(samples_dfITS.R$Replicate)
#------define the row names from the otu column- bacteria
asv_matITS.R <- asv_matITS.R %>%
  tibble::column_to_rownames("ASV") 
asv_matITS.R <- as.matrix(asv_matITS.R) 
#id matrix
tax_matITS.R <- tax_matITS.R %>% 
  tibble::column_to_rownames("ASV")
tax_matITS.R <- as.matrix(tax_matITS.R)
samples_dfITS.R <- samples_dfITS.R %>% 
  tibble::column_to_rownames("ID")
#------create phyloseq object
ASV.ITS = otu_table(asv_matITS.R, taxa_are_rows = TRUE)
TAX.ITS = tax_table(tax_matITS.R)
samples.16S = sample_data(samples_df16S.R)
df.bacteria <- phyloseq(ASV.16S, TAX.16S, samples.16S)
df.bacteria
#-------create phyoseq object just replicates 
df.replicates <- subset_samples(df.bacteria, Replicated == "Yes")
#Filter through samples 
df.replicates <- subset_taxa(df.replicates, Genus != "mitochondria_genus_unknown")
df.replicates <- subset_taxa(df.replicates, Genus != "Ralstonia")
df.replicates.fil = filter_taxa(df.replicates, function(x) sum(x > 5) > (0.25*length(x)), TRUE)
RA.replicates<- transform_sample_counts(df.replicates.fil, function(x) x / sum(x) )

#create taxa plots for replicate samples-favcet by rep
p <- plot_bar(RA.replicates, fill = "Genus", title = "Genus") + geom_point() 
+ theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

p + geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") + facet_wrap(~Replicate, scales= "free_x", nrow=1)

```


```{r}
#relative abundance to make S.cerevisae 
df.RA.ITS <- read_csv("RA.ITS.matrix.csv")
#Saccharomyces_species_unknown

ggplot(df.RA.ITS, aes(x=Type, y=Saccharomyces_species_unknown, fill=Type)) +
  geom_boxplot() +
  stat_compare_means(aes(label=paste0(..p.format..)), label.y = c(5,5)) +
  scale_y_continuous(limits=c(0,1))
  

```

