---
title: "SNPro Phyloseq"
author: "K Harrison"
date: "7/4/2022"
output: html_document
---

Load library
```{r setup, include=FALSE}
library(phyloseq)
library(devtools)
library(dplyr)
library(vegan)
library(ggplot2)
library(ape)
library(ggpubr)
library(readr)
library(devtools)
library(ggfortify)
library(ggforce)
library(ellipse)
library(data.table)
#library(decontam)
library(microViz)
install.packages(c("devtools", "RcppEigen", "RcppParallel", "Rtsne", "ggforce", "units"))
install.packages("remotes")
remotes::install_github("vmikk/metagMisc")
library(metagMisc)
library(FSA) #Dunn post hoc test for ANOVA
```


Load bacteria and yeast data
```{r}
setwd("~/Documents/Data/Probiotic SN Study/Packaged product analysis/R data")

#Bacteria
asv_mat16S<- read_csv("16S ASV SNPro Pro.csv") #ASV counts
tax_mat16S<- read_csv("16S TAXA SNPro.csv") #Taxa comparison
samples_df16S <- read_csv("16S SAMP SNPro Pro.csv") #Sample comparison
###set demo info to factors so that we can merge downstream
samples_df16S$Type <-as.factor(samples_df16S$Type)
samples_df16S$Claim <-as.factor(samples_df16S$Claim)
samples_df16S$Project <-as.factor(samples_df16S$Project)
samples_df16S$PSPECIES.LISTED <-as.factor(samples_df16S$PSPECIES.LISTED)
```

Define the row names from the otu column- bacteria
```{r}
#------define the row names from the otu column- bacteria
asv_mat16S <- asv_mat16S %>%
  tibble::column_to_rownames("ASV") 
asv_mat16S <- as.matrix(asv_mat16S) #idem matrix

tax_mat16S <- tax_mat16S %>% 
  tibble::column_to_rownames("ASV")
tax_mat16S <- as.matrix(tax_mat16S)

samples_df16S <- samples_df16S %>% 
  tibble::column_to_rownames("ID")
```


Create phyloseq objects
```{r}
#Bacteria
ASV.16S = otu_table(asv_mat16S, taxa_are_rows = TRUE)
TAX.16S = tax_table(tax_mat16S)
samples.16S = sample_data(samples_df16S)
df.bacteria <- phyloseq(ASV.16S, TAX.16S, samples.16S)
df.bacteria
```

DECONTAM
```{r}
df <- as.data.frame(sample_data(df.bacteria))
df$LibrarySize <- sample_sums(df.bacteria) #number of reads
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df)) #assign number to order
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Sample_or_Control)) + geom_point()
head(df$LibrarySize)

df.bacteria.ord <- ordinate(df.bacteria, "PCoA", "bray")
plot_ordination(df.bacteria, df.bacteria.ord, type="samples", color="Sample_or_Control")

sample_data(df.bacteria)$is.neg <- sample_data(df.bacteria)$Sample_or_Control == "Control Sample"
contamdf.prev.05 <- isContaminant(df.bacteria, method="prevalence", neg="is.neg", threshold=0.5)
table(contamdf.prev.05$contaminant)

row_indices <- which(contamdf.prev.05$contaminant)
```




Filter thru bacteria and create subset for Package Product Exp
```{r}
df.bacteria@sam_data$LibrarySize <- sample_sums(df.bacteria) #adding read abundance to phyloseq object
df.bacteria <- subset_taxa(df.bacteria, Genus != "mitochondria_genus_unknown")
df.bacteria <- subset_taxa(df.bacteria, Genus != "Ralstonia")
df.bacteria <- subset_samples(df.bacteria, NMR.Data != "No")
# Determine which genera are present
tax_table(df.bacteria)[, "Genus"]
#generate random tree and merge into phyloseq object for ordinate calculations
random_tree = rtree(ntaxa(df.bacteria), rooted=TRUE, tip.label=taxa_names(df.bacteria))
#plot(random_tree)
df.bacteria = merge_phyloseq(df.bacteria, random_tree)
df.bacteria.prod <- subset_samples(df.bacteria, Project == "Package Product")
df.bacteria.prod@sam_data$LibrarySize <- sample_sums(df.bacteria.prod)
#tax_table(df.bacteria.prod)[, "Genus"]
#create a dataset for the packaged product analysis
bacteria.prod.matrix <- phyloseq_to_df(df.bacteria.prod, addtax = T, addtot = T, addmaxrank = F,
  sorting = "abundance")
bacteria.prod.matrix<- t(bacteria.prod.matrix)#transpose matrix

df.bacteria.fil = filter_taxa(df.bacteria, function(x) sum(x > 5) > (0.25*length(x)), TRUE) ##Remove taxa not seen more than 5 times in at least 20% of the samples. This protects against an OTU with small mean & trivially large C.V.
#df.bacteria <- tax_glom(df.bacteria.fil, taxrank = "Genus") #merge genus-level
#tax_table(df.bacteria.fil)[, "Genus"]
RA.bacteria<- transform_sample_counts(df.bacteria.fil, function(x) x / sum(x) )
tax_table(RA.bacteria)[, "Genus"]
tax_table(df.bacteria.fil)[, "Genus"]

bacteria.RA.matrix <- phyloseq_to_df(RA.bacteria, addtax = T, addtot = T, addmaxrank = F,
  sorting = "abundance")

write.csv(bacteria.RA.matrix, "RA.16S.matrix.csv")
#Create taxa plots via claim and type
#```{r}
#p1 <- plot_bar(RA.bacteria, fill = "Genus", title = "Genus") + facet_wrap(~Type, scales= "free_x", nrow=1) + geom_point() + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
#p1 = p1 + geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack")
#p2 <- plot_bar(RA.bacteria, fill = "Family", title = "") + facet_wrap(~Type, scales= "free_x", nrow=1) + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
#p2 = p2 + geom_bar(aes(color=Family, fill=Family), stat="identity", position="stack") #+ theme(legend.position = "none") 
#p3 <- plot_bar(RA.bacteria, fill = "Genus", title = "") + facet_wrap(~Claim, scales= "free_x", nrow=1) + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
#p3 = p3 + geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") #+ theme(legend.position = "none") 
#p4 <- plot_bar(RA.bacteria, fill = "Family", title = "") + facet_wrap(~Claim, scales= "free_x", nrow=1) + theme(axis.ticks.x = element_blank(),axis.text.x = element_blank())
#p4 = p4 + geom_bar(aes(color=Family, fill=Family), stat="identity", position="stack") #+ theme(legend.position = "none") 

#ggarrange(p2,p4, nrow = 1,ncol=2,common.legend = TRUE, legend="right")
```


Make Taxa plots for samples with LOW and HIGH total CFU
```{r}
#create a dataset with low CFU count
df.low <- subset_samples(df.bacteria, CFU.COUNT=="<1000")
df.low.fil = filter_taxa(df.low, function(x) sum(x > 5) > (0.20*length(x)), TRUE)
RA.low<- transform_sample_counts(df.low.fil, function(x) x / sum(x) )

p.low <- plot_bar(RA.low, fill = "Genus", title = "Genus") + geom_point() + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())

p.low + geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") + facet_wrap(~Claim, scales= "free_x", nrow=1)
tax_table(df.low.fil)[, "Genus"]

df.hi <- subset_samples(df.bacteria,CFU.COUNT==">1000")
df.hi.fil = filter_taxa(df.hi, function(x) sum(x > 5) > (0.20*length(x)), TRUE)
RA.hi<- transform_sample_counts(df.hi.fil, function(x) x / sum(x) )
tax_table(df.hi.fil)[, "Genus"]
p.hi <- plot_bar(RA.hi, fill = "Genus", title = "Genus") + geom_point() + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
#Manually removed ASV: 5, 10, 83, 50, 57, 24, 96, 86, 53, 9

```


Make Taxa plots for samples with lactobacillus and bacillus listed
```{r}
#create a dataset with lactobacillus listed as ingredient
df.lact<- subset_samples(df.bacteria, LACTO.LISTED=="Y")
df.lact.fil = filter_taxa(df.lact, function(x) sum(x > 5) > (0.75*length(x)), TRUE)
RA.lact<- transform_sample_counts(df.low.fil, function(x) x / sum(x) )
p.lact <- plot_bar(RA.lact, fill = "Genus", title = "Genus") + geom_point() + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
p.lact + geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") 
#tax_table(df.low.fil)[, "Genus"]
#create a dataset with bacillus listed as ingredient
df.bacillus<- subset_samples(df.bacteria, BACILLUS.LISTED=="Y")
df.bacillus.fil = filter_taxa(df.bacillus, function(x) sum(x > 5) > (0.75*length(x)), TRUE)
RA.bacillus<- transform_sample_counts(df.bacillus.fil, function(x) x / sum(x) )
p.bacillus <- plot_bar(RA.bacillus, fill = "Genus", title = "Genus") + geom_point() + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
p.bacillus + geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") 

#Manually removed ASV: 5, 10, 83, 50, 57, 24, 96, 86, 53, 9

```


```
Alpha diversity comparison- btw types and claims 
reference: https://micca.readthedocs.io/en/latest/phyloseq.html
```{r}
#Measure Alpha diversity measure- Shannon
df.bacteria.shannon <- estimate_richness(df.bacteria.fil, measures="Shannon")
#Measure Alpha diversity measure- Chao1
df.bacteria.chao1 <- estimate_richness(df.bacteria.fil, measures="Chao1")

#Plot alpha diversity by Type, Claim, PStrain.Listed
p.alpha.claim <- plot_richness(df.bacteria.fil, "Claim", measures=c("Chao1", "Shannon"), color = "Claim") 

p.alpha.type <- plot_richness(df.bacteria.fil, "Type", measures=c("Chao1", "Shannon"), color = "Type") 
p.alpha.type + geom_boxplot(alpha=0.3)
p.alpha.probiotic <- plot_richness(df.bacteria.fil, "PSPECIES.LISTED", measures=c("Chao1", "Shannon"), color = "PSPECIES.LISTED") 
p.alpha.probiotic + geom_boxplot(alpha=0.3)

aov()

#Pairwise Wilcoxon rank-sum test for Beta Diversity
pairwise.wilcox.test(df.bacteria.chao1$Chao1,sample_data(df.bacteria.fil)$Type)
#              Hard Soft
#Soft          0.27 -   
#Soft-Aberrant 0.44 0.33
pairwise.wilcox.test(df.bacteria.chao1$Chao1,sample_data(df.bacteria.fil)$Claim)
#         live  none 
#none      0.053 -    
#probiotic 0.052 0.224
#chi-square for PSpecies.Listed.....

pairwise.wilcox.test(df.bacteria.shannon$Shannon,sample_data(df.bacteria.fil)$Type)
#              Hard Soft
#Soft          0.18 -   
#Soft-Aberrant 0.67 0.10
pairwise.wilcox.test(df.bacteria.shannon$Shannon,sample_data(df.bacteria.fil)$Claim)
#          live none
#none      0.85 -   
#probiotic 0.20 0.85
```
p.alpha.type + geom_boxplot(data=p.alpha.probiotic$data, aes(x=Type, y=value, fill=Type), alpha=0.3) + theme_bw()

Ordination plots
```{r}
bacteria.ordan <- ordinate(df.bacteria.fil, "NMDS", "bray")
p1 = plot_ordination(df.bacteria.fil, bacteria.ordan, type="taxa", color="Family", title="taxa")
print(p1)
```


Ordinate plots- PIE CHART
```{r}
df.bacteria.fil <- tax_fix(df.bacteria.fil)
df.bacteria.fil <- phyloseq_validate(df.bacteria.fil, remove_undetected = TRUE)
df.bacteria.fil %>% 
  tax_transform(trans = "clr", rank = "Genus")


df.bacteria.fil %>% 
  tax_transform("clr", rank = "Genus") %>% 
  # when no distance matrix or constraints are supplied, PCA is the default/auto ordination method
  ord_calc() %>% 
  ord_plot_iris(tax_level = "Genus", anno_colour = "Claim"
)

#scree
#distanceMethodList
DistUF = distance(df.bacteria.fil, method = "bray")
ordUF = ordinate(df.bacteria.fil, method = "PCoA", distance = DistUF)
plot_scree(ordUF, "Scree Plot: Weighted UniFrac MDS") + theme_bw()

adonis2(DistUF ~ sample_data(df.bacteria.fil)$Type) 
#p-value = 0.012*
adonis2(DistUF ~ sample_data(df.bacteria.fil)$Claim) 
#p-value = 0.041 *
adonis2(DistUF ~ sample_data(df.bacteria.fil)$PSPECIES.LISTED)
#0.002 **
#

plot_ordination(df.bacteria.fil, ordinate(df.bacteria.fil, method='PCoA', dist = 'bray'), type='split', 
                color='Type')
ord <- ordinate(df.bacteria.fil, method='PCoA')
p.ord <- plot_ordination(df.bacteria.fil, ord, type='sample', 
                color="Type")
#t-distribution and normal distribution         
p.ord + 
  stat_ellipse(type = "norm", linetype = 2) +
  stat_ellipse(type = "t") + 
  theme_bw()
```

p + 
  stat_ellipse(type = "norm", linetype = 2) +
  stat_ellipse(type = "t") + 
  theme_bw()



df.bacteria.fil %>% 
  tax_transform("clr", rank = "Genus") %>% 
  # when no distance matrix or constraints are supplied, PCA is the default/auto ordination method
  ord_calc() %>% 
  ord_plot_iris(tax_level = "Genus",  anno_colour = "Type"
)


```{r}
df.type <- subset_samples(df.bacteria.fil, PSPECIES.LISTED=="NO")
df.type <- subset_taxa(df.type, Genus != "mitochondria_genus_unknown")
df.type <- tax_fix(df.type)
df.type <- phyloseq_validate(df.type, remove_undetected = TRUE)

df.type %>% 
  tax_transform(trans = "clr", rank = "Genus")

df.type %>% 
  tax_transform("clr", rank = "Genus") %>% 
  # when no distance matrix or constraints are supplied, PCA is the default/auto ordination method
  ord_calc() %>% 
  ord_plot_iris(tax_level = "Genus", ord_plot = "above", anno_colour = "Type"
)

df.prob<- subset_samples(df.bacteria.fil, PSPECIES.LISTED=="YES")
df.prob <- subset_taxa(df.prob, Genus != "mitochondria_genus_unknown")
df.prob <- tax_fix(df.prob)
df.prob <- phyloseq_validate(df.prob, remove_undetected = TRUE)

df.prob %>% 
  tax_transform(trans = "clr", rank = "Genus")

df.prob %>% 
  tax_transform("clr", rank = "Genus") %>% 
  # when no distance matrix or constraints are supplied, PCA is the default/auto ordination method
  ord_calc() %>% 
  ord_plot_iris(tax_level = "Genus", ord_plot = "above", anno_colour = "Type"
)
```
df.type %>% 
  tax_transform("clr", rank = "Genus") %>% 
  # when no distance matrix or constraints are supplied, PCA is the default/auto ordination method
  ord_calc() %>% 
  ord_plot_iris(tax_level = "Genus"
)

df.prob %>% 
  tax_transform("clr", rank = "Genus") %>% 
  # when no distance matrix or constraints are supplied, PCA is the default/auto ordination method
  ord_calc() %>% 
  ord_plot_iris(tax_level = "Genus",anno_colour = "ADDED.SPECIES",  ord_plot = "above"
)
  
Make a dataframe for just samples with Lacto and Bacillus clalims 

```{r}
df.lacto <- subset_samples(df.bacteria, LACTO.LISTED == "Y")
df.lacto.fil = filter_taxa(df.lacto, function(x) sum(x > 5) > (0.25*length(x)), TRUE) ##Remove taxa not seen more than 5 times in at least 20% of the samples. This protects against an OTU with small mean & trivially large C.V.
RA.lacto<- transform_sample_counts(df.lacto.fil, function(x) x / sum(x) )

pL <- plot_bar(RA.lacto, fill = "Genus", title = "Lactobacillus Listed") 
  # + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
pL = pL + geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") #+ theme(legend.position = "none") 


df.bacillus <- subset_samples(df.bacteria, BACILLUS.LISTED == "Y")
df.bacillus.fil = filter_taxa(df.bacillus, function(x) sum(x > 5) > (0.25*length(x)), TRUE) ##Remove taxa not seen more than 5 times in at least 20% of the samples. This protects against an OTU with small mean & trivially large C.V.
RA.bacillus<- transform_sample_counts(df.bacillus.fil, function(x) x / sum(x) )

pB <- plot_bar(RA.bacillus, fill = "Genus", title = "Bacillus Listed") 
  # + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
pB = pB + geom_bar(aes(color=Genus, fill=Genus), stat="identity", position="stack") #+ theme(legend.position = "none") 

ggarrange(pL,pB, nrow = 1,ncol=2,common.legend = TRUE)
```


```{r}
df.bacteria <- phyloseq_to_df(df.bacteria.fil, addtax = T, addtot = F, addmaxrank = F,
  sorting = "abundance")
write.csv(df.bacteria,"~/Documents/Data/Probiotic SN Study/Packaged product analysis/R data/dfbacteria.csv", row.names = TRUE)
```

Distribution of lacto and bacillus reads by Type
```{r}
df <- as.data.frame(df.bacteria.fil@otu_table)
#transform data frame
df.reads.bacteria <- t(df) 
#ASV002 bacillus ASV003 Lactobacillus
#Add Type variable as a column to df.reads.bacteria
#df.reads.LAB <- df.reads.bacteria[,c(4,7)]

df.reads.LAB<- read_csv("Lac.Bac.Reads.csv") #ASV counts

#make histogram of bacilllus and lactobacillus distribution
bacillus.hist <- ggplot(df.reads.LAB, aes(x=`Bacillus Log(Reads)`, color=Type, fill=Type)) +
   geom_histogram(position="identity", alpha=0.7)  + geom_density(alpha=0.25)
   
bacillus.hist.l <- ggplot(df.reads.LAB, aes(x=`Bacillus Log(Reads)`, color=BACILLUS.LISTED, fill=BACILLUS.LISTED)) +
   geom_histogram(position="identity", alpha=0.9)  + geom_density(alpha=0.5) + theme_bw()
   
lactobacillus.hist <- ggplot(df.reads.LAB, aes(x=`Lactobacillus Log(Reads)`, color=Type, fill=Type)) +
   geom_histogram(position="identity", alpha=0.7)  + geom_density(alpha=0.25)

lactobacillus.hist.l <- ggplot(df.reads.LAB, aes(x=`Lactobacillus Log(Reads)`, color=LACTO.LISTED, fill=LACTO.LISTED)) +
   geom_histogram(position="identity", alpha=0.9)  + geom_density(alpha=0.5) + theme_bw() +
  scale_color_manual(values = c("gold1", "darkgrey")) +
  scale_fill_manual(values = c("gold1", "darkgrey"))



#ANOVA tests for distribution
kruskal.test(`ASV002`~ Type, data = df.reads.LAB) 
#chi-squared = 16.422, df = 2, p-value = 0.0002717
kruskal.test(`ASV003`~ Type, data = df.reads.LAB) 

#ANOVA tests for distribution
kruskal.test(`ASV002`~ BACILLUS.LISTED, data = df.reads.LAB) 
#chi-squared = 16.422, df = 2, p-value = 0.0002717
kruskal.test(`ASV003`~ LACTO.LISTED, data = df.reads.LAB) 
```
library(patchwork)
lactobacillus.hist + scale_color_brewer(palette="Dark2") + scale_fill_brewer(palette="Dark2")
bacillus.hist.l|lactobacillus.hist.l

```{r}
RA.ANOVA<- read_csv("RA.16S.NMR.csv") #ASV counts

kruskal.test(`Gluconobacter`~ Type, data = RA_16S_NMR) 
#p-value =0.01775
dunnTest(`Gluconobacter`~ Type, data = RA_16S_NMR, method="bh") 
#Hard-Soft p = 0.02125937

kruskal.test(`Bacillus`~ Type, data = RA_16S_NMR) 
#p-value = 0.02697
dunnTest(`Bacillus`~ Type, data = RA.ANOVA, method="bh") 
#Hard-Soft p = 0.04962812

kruskal.test(`Lactobacillus`~ Type, data = RA.ANOVA) 
#p-value = 0.04015
dunnTest(`Lactobacillus`~ Type, data = RA.ANOVA, method="bh") 
#Hard-Soft p = 0.03383108

kruskal.test(`Komagateibacter`~ Type, data = RA.ANOVA) 
#p-value = 0.2797

kruskal.test(`Streptococcus`~ Type, data = RA.ANOVA) 
#p-value = 0.003987
dunnTest(`Streptococcus`~ Type, data = RA.ANOVA, method="bh") 
#Hard-Soft-Aberrant p = 0.006214781; Hard-Soft p-value = 0.009481806

kruskal.test(`Acetobacter`~ Type, data = RA.ANOVA) 
#p-value = 0.5697

kruskal.test(`Acinetobacter`~ Type, data = RA.ANOVA) 
#p-value = 0.01785
dunnTest(`Acinetobacter`~ Type, data = RA.ANOVA, method="bh") 
#Hard-Soft p = 0.01478114

kruskal.test(`Acetobacteraceae_genus_unknown`~ Type, data = RA.ANOVA) 
#p-value = 0.05766

kruskal.test(`Enterobacteriaceae_genus_unknown`~ Type, data = RA.ANOVA) 
#p-value = 0.5067

kruskal.test(`Chryseobacterium`~ Type, data = RA.ANOVA) 
#p-value = 0.03035
dunnTest(`Chryseobacterium`~ Type, data = RA.ANOVA, method="bh") 
#Hard-Soft p = 0.02460663

kruskal.test(`Sphingomonas`~ Type, data = RA.ANOVA) 
#p-value = 0.04861
dunnTest(`Sphingomonas`~ Type, data = RA.ANOVA, method="bh") 
#Hard-Soft p = 0.05349564

kruskal.test(`Sphingobium"`~ Type, data = RA.ANOVA) 
#p-value = 0.04861
dunnTest(`Sphingobium`~ Type, data = RA.ANOVA, method="bh") 
#Hard-Soft p = 0.05349564

kruskal.test(`Sphingobium"`~ Type, data = RA.ANOVA) 
#p-value = 0.00854
dunnTest(`Sphingobium"`~ Type, data = RA.ANOVA, method="bh") 
#Hard-Soft p = 0.00548857; S-SA p-vale = 0.029977080

kruskal.test(`Pedobacter`~ Type, data = RA.ANOVA) 
#p-value = 0.05898
```

Looking at replicate samples 
```{r}
#Bacteria
asv_mat16S.R<- read_csv("16S ASV SNPro.csv") #ASV counts
tax_mat16S.R<- read_csv("16S TAXA SNPro.csv") #Taxa comparison
samples_df16S.R <- read_csv("16S SAMP SNPro.csv") #Sample comparison
###set demo info to factors so that we can merge downstream
samples_df16S.R$Replicated <-as.factor(samples_df16S.R$Replicated)
samples_df16S.R$Replicate <-as.factor(samples_df16S.R$Replicate)
#------define the row names from the otu column- bacteria
asv_mat16S.R <- asv_mat16S.R %>%
  tibble::column_to_rownames("ASV") 
asv_mat16S.R <- as.matrix(asv_mat16S.R) 
#id matrix
tax_mat16S.R <- tax_mat16S.R %>% 
  tibble::column_to_rownames("ASV")
tax_mat16S.R <- as.matrix(tax_mat16S.R)
samples_df16S.R <- samples_df16S.R %>% 
  tibble::column_to_rownames("ID")
#------create phyloseq object
ASV.16S = otu_table(asv_mat16S.R, taxa_are_rows = TRUE)
TAX.16S = tax_table(tax_mat16S.R)
samples.16S = sample_data(samples_df16S.R)
df.bacteria <- phyloseq(ASV.16S, TAX.16S, samples.16S)
df.bacteria
#-------create phyoseq object just replicates 
df.replicates <- subset_samples(df.bacteria, Replicated == "Yes")
#Filter through samples 
df.replicates <- subset_taxa(df.replicates, Genus != "mitochondria_genus_unknown")
df.replicates <- subset_taxa(df.replicates, Genus != "Ralstonia")
df.replicates.fil = filter_taxa(df.replicates, function(x) sum(x > 5) > (0.25*length(x)), TRUE)

#create taxa plots for replicate samples-favcet by rep

```

