---
title: "Packaged Product-SN-June 2022"
author: "K Harrison"
date: "7/1/2022"
output: html_document
---

Evaluate NMR and CFU data from packaged probiotic study 
-----

Load libraries
```{r}
library(dplyr)
library(plyr)
library(vegan)
library(ape)
library(devtools)
library(ggplot2)
library(ggfortify)
library(ggrepel) #repel loading labels
library(vtable) #creates tables of summary stats
library(gridExtra)
library(FSA) #Dunn post hoc test for ANOVA
library(PCAtools) #for making PCA plots 
library(ggforce)
library(ellipse)
library(readr)
library(tidyverse)
library(tidyr)
library(ggpubr) #ggarrange
library(leaps)
library(corrplot)
library(car)
library(GGally)
library(viridis)
library(gridExtra) #MANOVA
library(MASS) #LDA post-hoc MANOVA
library(patchwork)
```

Load data
```{r}
#set working directory
setwd("~/Documents/Data/Probiotic SN Study/Packaged product analysis/R data")

#import data 
df.snpro <- read_csv("SNProN2.csv") #original dataset containing demograohic, cfu, logcfu, and NMR data- no samples removed
#df.snpro.legal <- read_csv("SNProR.legal.csv") #original dataset containing demograohic, cfu, logcfu, and NMR data- soft samples <0.5% have been removed
#View(df)
#summary(df)
#df.readcounts <-  read_csv("Lac.Bac.Reads.csv")
#df.readcounts <- df.readcounts[,-1]
#make first column row names
#------------------------------------adjust SNPro dataset (CFU-Claim)
df.adj <- df.snpro[-c(4,28,26,24,22)] #remove non-log counts & categorical columnswith NA
#df.adj[is.na(df.adj)] <- 0 #convert NA to 0
df.adj$SNPRO <- as.character(df.adj$SNPRO) #convert IDs to characters

df.adj$'Total acid(g L-1)' <- df.adj$`Acetic acid(g L-1)` + df.adj$`Lactic acid(g L-1)` + df.adj$`Gluconic acid(g L-1)` #calculate total acidity
df.adj$'Total sugar(g L-1)' <- df.adj$`Glucose(g L-1)` + df.adj$`Fructose(g L-1)` + df.adj$`Sucrose(g L-1)` #calculate TOTAL SUGAR
df.nmr <- subset(df.adj, NMR.Data == "Yes")
df.nmr.NOethanol <- df.nmr[c(2,4,10,11,14:20,27,28)] #DF with all metabolites but etoh
df.nmr.etoh<- df.nmr[c(2,4,10,11,12,14:20,27,28)] #DF with all metabolites
#ethanol removed
df.cfu <- df.adj[c(2,4,7,9,21:24,26)] #just log cfu values, claim, strainlisted, and type
df.cfu.type <- df.nmr[c(2,4,7,9,21:24,26)] #just log cfu values, claim, strainlisted, and type

#------------------------------------adjust SNPro Legal dataset (NMR)
#df.adj.l <- df.snpro.legal[-c(1,3,4,5,6,7,11,26:30)] #remove non-log counts & categorical columnswith NA
#df.adj.l[is.na(df.adj.l)] <- 0 #convert NA to 0
#df.adj.l$SNPRO <- as.character(df.adj.l$SNPRO) #convert IDs to characters
#-------------------------------- CFU dataset

CFU.all <- read_csv("CFU.All.csv") #reformated logCFU data to make boxplots with multiple levels
CFU.all[is.na(CFU.all)] <- 0

SPC.all <- read_csv("SPC.counts.csv") #reformated logCFU counts to make boxplots with levels across all media types and median cfu ml-1


#---------------------------------- Probiotic subset
df.cfu.prob <- subset(df.cfu, Claim == "probiotic")
#----------------------------Summary stats- demographics 
df.demo <- df.nmr[c(1,2,4:6)]
#df.demo.l <- df.adj.l[c(1:4)]
df.demo.cfu <- df.adj[c(4,21:25)]
st(df.demo)
df.nmr.demo <- df.nmr[-c(2)]
st(df.nmr, group = 'Type', group.test = FALSE, digits = 2, fixed.digits = TRUE)
st(df.demo.cfu, group = 'Claim', group.test = FALSE, digits = 2, fixed.digits = TRUE)
```

Descriptive statistics 
```{r}
#make hisograms to determine normality 
#hist.data.frame(df.adj) #nothing has a normal distribution

#make 2x2 panel of boxplots for CFU/ml by Claim
#attach(df.adj)
#par(mfrow=c(2,2))
#boxplot(logM13 ~ Claim)
#boxplot(logMRS ~ Claim)
#boxplot(logWLN ~ Claim)
#boxplot(logNA ~ Claim)
```

Create violin plots for overall count by claim- including median CFU
```{r}
#Re-order groups: change the order of SPC.all$Media
SPC.all$Media <- factor(SPC.all$Media, levels = c("NAm", "M13", "MRS", "WLN", "Median"), order = TRUE)

SPC.all.boxplot <- ggplot(SPC.all, aes(x= Media, y=`log(CFU ml-1)`, fill=Media)) +
  geom_violin()+
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="BuPu")
```

SPC.all.boxplot

WLN
Create violin plots of SPC by claim
```{r}
#remove samples with "none" as claim
#make violin plots for CFU/ml distribution across different media types

M13 <-  ggplot(df.adj, aes(x= Claim, y=`logM13`, fill=Claim)) +
  geom_violin()+
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu")

MRS <- ggplot(df.adj, aes(x= Claim, y=`logMRS`, fill=Claim)) +
  geom_violin()+
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu")

WLN <- ggplot(df.adj, aes(x= Claim, y=`logWLN`, fill=Claim)) +
  geom_violin()+
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu")
NAm <- ggplot(df.adj, aes(x= Claim, y=`logNA`, fill=Claim)) +
  geom_violin()+
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu")
#arrange plots into one plot with multiple panels
ggarrange(M13, MRS, NAm, WLN, nrow = 4, common.legend = TRUE, legend="bottom") 


# Install and load the ggstatsplot package to include mean
install.packages("ggplot2", dependencies = TRUE)
install.packages("ggstatsplot")
library(ggstatsplot)

# Modify the code for each plot to include statistics
M13 <- ggplot(df.adj, aes(x= Claim, y=`logM13`, fill=Claim)) +
  geom_violin()+
  geom_boxplot(width=0.1) + 
  geom_jitter(alpha=0.8, size=1, width=0.1) + 
  scale_fill_brewer(palette="RdPu")

MRS <- ggplot(df.adj, aes(x= Claim, y=`logMRS`, fill=Claim)) +
  geom_violin()+
  geom_boxplot(width=0.1) + 
  geom_jitter(alpha=0.8, size=1, width=0.1) + 
  scale_fill_brewer(palette="RdPu")

WLN <- ggplot(df.adj, aes(x= Claim, y=`logWLN`, fill=Claim)) +
  geom_violin()+
  geom_boxplot(width=0.1) + 
  geom_jitter(alpha=0.8, size=1, width=0.1) + 
  scale_fill_brewer(palette="RdPu")

NAm <- ggplot(df.adj, aes(x= Claim, y=`logNA`, fill=Claim)) +
  geom_violin()+
  geom_boxplot(width=0.1) + 
  geom_jitter(alpha=0.8, size=1, width=0.1) + 
  scale_fill_brewer(palette="RdPu")

# Arrange the plots into one plot with multiple panels
ggarrange(M13, MRS, NAm, WLN, nrow = 4, common.legend = TRUE, legend="bottom")

#create dataframe with only "probiotic" and "probitoic and live" by probiotic ingredient 

M13.s <- ggplot(df.adj, aes(x= Organism.Listed, y=`logM13`, fill=Organism.Listed)) +
  geom_violin()+
  geom_jitter(alpha=0.8, size=1, width=0.1) + 
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") + scale_x_discrete(guide = guide_axis(check.overlap = TRUE))
MRS.s <- ggplot(df.adj, aes(x= Organism.Listed, y=`logMRS`, fill=Organism.Listed)) +
  geom_violin()+
  geom_jitter(alpha=0.8, size=1, width=0.1) + 
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") + scale_x_discrete(guide = guide_axis(check.overlap = TRUE)) 
WLN.s <- ggplot(df.adj, aes(x= Organism.Listed, y=`logWLN`, fill=Organism.Listed)) +
  geom_violin()+
  geom_jitter(alpha=0.8, size=1, width=0.1) + 
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") + scale_x_discrete(guide = guide_axis(check.overlap = TRUE))
NAm.s <- ggplot(df.adj, aes(x= Organism.Listed, y=`logNA`, fill=Organism.Listed)) +
  geom_violin()+
  geom_jitter(alpha=0.8, size=1, width=0.1) + 
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") + scale_x_discrete(guide = guide_axis(check.overlap = TRUE))
Max.s <- ggplot(df.adj, aes(x= Organism.Listed, y=Max.cfu, fill=Organism.Listed)) +
  geom_violin()+
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") + scale_x_discrete(guide = guide_axis(check.overlap = TRUE)) + theme_bw()
#arrange plots into one plot with multiple panels
Organism.CFU.box <- ggarrange(MRS.s,Max.s,nrow = 2, ncol=2, common.legend = TRUE, legend="bottom") 

```

ggarrange(M13, MRS, NAm, WLN, nrow = 4,common.legend = TRUE, legend="bottom") 
ggarrange(M13.s, MRS.s, NAm.s, WLN.s, nrow = 4, common.legend = TRUE, legend="bottom") 
ggarrange(M13.t, MRS.t, NAm.t, WLN.t, nrow = 4, common.legend = TRUE, legend="bottom") 
organism <- ggarrange(MRS.s, Max.s, nrow = 1, ncol=2, common.legend = TRUE, legend="bottom") 


```{r}
#make violin plots for CFU/ml distribution across different media types

M13.t <- ggplot(df.nmr, aes(x= Type, y=`logM13`, fill=Type)) +
  geom_violin()+
  geom_jitter(alpha=0.8, size=1, width=0.1) + 
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") + theme_bw() + theme(legend.position="bottom")
MRS.t <- ggplot(df.nmr, aes(x= Type, y=`logMRS`, fill=Type)) +
  geom_violin()+
  geom_jitter(alpha=0.8, size=1, width=0.1) + 
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") + theme_bw()
WLN.t <- ggplot(df.adj, aes(x= Type, y=`logWLN`, fill=Type)) +
  geom_violin()+
  geom_jitter(alpha=0.8, size=1, width=0.1) + 
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") 
NAm.t <- ggplot(df.nmr, aes(x= Type, y=`logNA`, fill=Type)) +
  geom_violin()+
  geom_jitter(alpha=0.8, size=1, width=0.1) + 
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") + theme(legend.position = "none")
#arrange plots into one plot with multiple panels
ggarrange(M13.t, MRS.t, NAm.t, WLN.t, nrow = 4, common.legend = TRUE, legend="bottom") 



```
(hist.cfu.claim + hist.cfu.type)/(M13.t + p)

```{r}
data <- df.nmr[c(1,3,4,5,7,8,9,12,13)]

# Plot
ggparcoord(data,
    columns = 2:9, groupColumn = 1,showPoints = TRUE, scale="center")
+ 
  scale_color_viridis(discrete=TRUE)
```

Histogram by group
```{r}
p1 = ggplot(df.nmr, aes(x=`Gluconic acid(g/L)`, color=Type, fill=Type)) +
   geom_histogram(position="identity", alpha=0.5)  + geom_density(alpha=0.25)
p2 = ggplot(df.nmr, aes(x=`Total Sugar(g/L)`, color=Type,fill=Type)) +
   geom_histogram(position="identity", alpha=0.5) + geom_density(alpha=0.25)
p5 = ggplot(df.nmr, aes(x=`Total Acid(g/L)`, color=Type,fill=Type)) +
   geom_histogram(position="identity", alpha=0.5) + geom_density(alpha=0.25)
p6 = ggplot(df.nmr, aes(x=`Acetic acid(g/L)`, color=Type,fill=Type)) +
   geom_histogram(position="identity", alpha=0.5) + geom_density(alpha=0.25)

#Distribution of ethanol across Types
p6 = ggplot(df.nmr, aes(x=`Ethanol(g L-1)`, color=Type,fill=Type)) +
   geom_histogram(position="identity", alpha=0.7) + geom_density(alpha=0.25)


ggarrange(p1,p6, p4, nrow = 1, common.legend = TRUE, legend = "bottom")
```
p6 = p6 + theme_bw() 

Histograms for max distribution
```{r}

max.type <- ggplot(df.cfu.type, aes(x=Max.cfu, color=Type, fill=Type)) +
   geom_histogram(position="identity", alpha=0.7)  + geom_density(alpha=0.25)

hist.cfu.type<- max.type + theme_bw() + scale_fill_brewer(palette="Dark2") + scale_color_brewer(palette="Dark2") + theme(legend.position = "right")
hist.cfu.type

max.claim <- ggplot(df.cfu, aes(x=Max.cfu, color=Claim, fill=Claim)) +
   geom_histogram(position="identity", alpha=0.7)  + geom_density(alpha=0.25)

hist.cfu.claim<- max.claim + theme_bw() + scale_fill_brewer(palette="Dark2") + scale_color_brewer(palette="Dark2") + theme(legend.position="bottom")
```
ggarrange(hist.cfu.type, hist.cfu.claim, nrow = 1, ncol = 2, common.legend = FALSE, legend = "bottom")


ANOVA- comparison between soft,soft-volatile and hard 
```{r}
kruskal.test(`Glucose(g L-1)`~ Type, data = df.nmr) #chi-squared = 16.422, df = 2, p-value = 0.0002717
kruskal.test(`Fructose(g L-1)`~ Type, data = df.nmr) #chi-squared 13.276, df = 2, p-value = 0.00131
kruskal.test(`Sucrose(g L-1)`~ Type, data = df.nmr) #chi-squared = 22.829, df = 2, p-value = 1.104e-05
kruskal.test(`Total sugar(g L-1)`~ Type, data = df.nmr) #chi-squared = 19.179, df = 2, p-value = 6.843e-05
kruskal.test(`Acetic acid(g L-1)`~ Type, data = df.nmr) #chi-squared = 8.8737, df = 2, p-value = 0.01183
kruskal.test(`Gluconic acid(g L-1)`~ Type, data = df.nmr) #chi-squared = 8.1374, df = 2, p-value = 0.0171
kruskal.test(`Lactic acid(g L-1)`~ Type, data = df.nmr)#chi-squared = 3.4333, df = 2, p-value = 0.1797
kruskal.test(`Total acid(g L-1)`~ Type, data = df.nmr)#chi-squared = 3.7664, df = 2, p-value = 0.1521
kruskal.test(`Erythritol(g L-1)`~ Type, data = df.nmr)#chi-squared = 7.0955, df = 2, p-value = 0.02879
kruskal.test(`Ethanol(g L-1)`~ Type, data = df.nmr)#chi-squared = 7.0955, df = 2, p-value = 0.02879
kruskal.test(`Glycerol(g L-1)`~ Type, data = df.nmr)#chi-squared = 7.0955, df = 2, p-value = 1.814e-05
kruskal.test(`Isobutyrate(mg L-1)`~ Type, data = df.nmr)#chi-squared = 7.0955, df = 2, p-value = 0.04035
```
Post-hoc Dunn Test for Kruskal-Wallis rank sum test ~Type
```{r}
dunnTest(`Glucose(g L-1)`~ Type, data = df.nmr, method="bh") #Hard - Soft; Hard - Soft-Aberrant
dunnTest(`Acetic acid(g L-1)`~ Type, data = df.nmr, method="bh") #Hard - Soft; Hard - Soft-Aberrant
dunnTest(`Ethanol(g L-1)`~ Type, data = df.nmr, method="bh") #Hard - Soft; Hard - Soft-Aberrant, Soft - Soft-Aberrant
dunnTest(`Fructose(g L-1)`~ Type, data = df.nmr, method="bh") #Hard - Soft; Hard - Soft-Aberrant
dunnTest(`Glycerol(g L-1)`~ Type, data = df.nmr, method="bh") #Hard - Soft; Hard - Soft-Aberrant; Soft - Soft-Aberrant
dunnTest(`Sucrose(g L-1)`~ Type, data = df.nmr, method="bh") #Hard - Soft; Hard - Soft-Aberrant, Soft - Soft-Aberrant
dunnTest(`Total sugar(g L-1)`~ Type, data = df.nmr, method="bh") #Hard - Soft; Hard - Soft-Aberrant, Soft - Soft-Aberrant
```


One-way ANOVA by Claim, Strain.Listed, and Type
```{r}


#nonparametric Kruskal-Wallis ANOVA
kruskal.test(`logMRS` ~ Claim, data = df.cfu) #p-value = 0.501
kruskal.test(`logM13` ~ Claim, data = df.cfu) #p-value = 0.06999
kruskal.test(`logNA` ~ Claim, data = df.cfu) #p-value = 0.1545
kruskal.test(`logWLN` ~ Claim, data = df.cfu) #p-value = 0.4175
kruskal.test(Max.cfu ~ Claim, data = df.cfu) #p-value = 0.07491

#nonparametric Kruskal-Wallis ANOVA
kruskal.test(`logMRS` ~ Type, data = df.cfu.type) #p-value = 0.08
kruskal.test(`logM13` ~ Type, data = df.cfu.type)  #p-value = 0.012
dunnTest(`logM13` ~ Type, data = df.cfu.type, method="bh") #p= 0.004 H-SA, S-SA
kruskal.test(`logNA` ~ Type, data = df.cfu.type)  #p-value = 0.085
kruskal.test(`logWLN` ~ Type, data = df.cfu.type)  #p-value = 0.06
dunnTest(`logWLN` ~ Type, data = df.cfu.type, method="bh") #p= 0.03 H-SA
kruskal.test(Max.cfu ~ Type, data = df.cfu.type)  #p-value = 0.06
dunnTest(`Max.cfu` ~ Type, data = df.cfu.type, method="bh") 
```
`
ggarrange(MRS.s, Max.s, nrow = 1, ncol = 2, common.legend = TRUE, legend = "bottom")



```{r}
``{r}
#nonparametric Kruskal-Wallis ANOVA
kruskal.test(`logMRS` ~ Organism.Listed, data = df.cfu) #Kruskal-Wallis hi-squared = 4.7872, df = 1, p-value = 0.0004179
kruskal.test(`logM13` ~ Organism.Listed, data = df.cfu) #Kruskal-Wallis chi-squared = 1.6969, df = 1, p-value =  0.2003
kruskal.test(`logNA` ~ Organism.Listed, data = df.cfu) #Kruskal-Wallis chi-squared = 0.080797, df = 1, p-value = 0.7127
kruskal.test(`logWLN` ~ Organism.Listed, data = df.cfu) #Kruskal-Wallis chi-squared = 2.148, df = 1, p-value = 0.1428
kruskal.test(Max.cfu ~ Organism.Listed, data = df.cfu) #Kruskal-Wallis chi-squared = 1.1266, df = p-value = 0.007331

#nonparametric Kruskal-Wallis ANOVA
kruskal.test(`logMRS` ~ Strain.Listed, data = df.cfu) #Kruskal-Wallis chi-squared = 1.8066, df = 1, p-value = 0.004056
kruskal.test(`logM13` ~ Strain.Listed, data = df.cfu) #df = 1,  p-value = 0.3976
kruskal.test(`logNA` ~ Strain.Listed, data = df.cfu) #Kruskal-Wallis chi-squared = 0.59788, df = 1, p-value = 0.1547
kruskal.test(`logWLN` ~ Strain.Listed, data = df.cfu) #Kruskal-Wallis chi-squared = 0.28378, df = 1, p-value = 0.2699
kruskal.test(Max.cfu ~ Strain.Listed, data = df.cfu) #Kruskal-Wallis chi-squared = 1.1266, df = 1, p-value = 0.007548

```
```







Make PCA plot by Type: Hard vs Soft vs Soft-aberrant
```{r}
#......for NMR data without Ethanol
View(df.nmr.NOethanol)
#create pca values for subsets
nmr.pca <- prcomp(df.nmr.NOethanol[c(3:13)], scale. = TRUE)
#plot biplot with ellipses (5% CI?)
autoplot(nmr.pca, data=df.nmr, colour="Type", loadings = TRUE, loadings.colour = 'brown',
         loadings.label.colour='brown', loadings.label = TRUE, loadings.label.size = 4,
         loadings.label.repel=TRUE) 
         +   ggforce::geom_mark_ellipse(aes(fill = Type,
                        color = Type))
                        
nmr.pca.etoh <- prcomp(df.nmr.etoh[c(3:14)], scale. = TRUE)
#plot biplot with ellipses (5% CI?)
autoplot(nmr.pca.etoh, data=df.nmr.etoh, colour="Type", loadings = TRUE, loadings.colour = 'brown',
         loadings.label.colour='brown', loadings.label = TRUE, loadings.label.size = 4,
         loadings.label.repel=TRUE) 
         +   ggforce::geom_mark_ellipse(aes(fill = Type,
                        color = Type))                        
                        
```
p7 = autoplot(nmr.pca, data=df.nmr, colour="Type", loadings = TRUE, loadings.colour = 'brown',
         loadings.label.colour='black', loadings.label = TRUE, loadings.label.size = 3,
         loadings.label.repel=TRUE) + theme_bw()

p7
Make PCA plot by Claim and Type: 

```{r}
cfu.claims <- df.claims[c(2,4,18:21)]
cfu.listed <- df.claims[c(6,18:21)]

#create pca values for subsets:Claims
cfu.claim.pca <- prcomp(cfu.claims[c(3:6)], scale. = TRUE)
#plot biplot with elliipses around oxygen levels 
autoplot(cfu.claim.pca, data=cfu.claims, colour="Claim", loadings = TRUE, loadings.colour = 'brown',
         loadings.label.colour='black', loadings.label = TRUE, loadings.label.size = 3,
         loadings.label.repel=TRUE) 

#create pca values for subsets:Strain.Listed
cfu.claim.pca <- prcomp(cfu.claims[c(3:6)], scale. = TRUE)
#plot biplot with elliipses around oxygen levels 
autoplot(cfu.claim.pca, data=df.claims, colour="Strain.Listed", loadings = TRUE, loadings.colour = 'brown',
         loadings.label.colour='black', loadings.label = TRUE, loadings.label.size = 3,
         loadings.label.repel=TRUE) 

#create pca values for subsets
cfu.type.pca <- prcomp(df.cfu[c(2:5)], scale. = TRUE)
#plot biplot with elliipses around oxygen levels 
autoplot(cfu.type.pca, data=df.cfu, colour="Type", loadings = TRUE, loadings.colour = 'brown',
         loadings.label.colour='black', loadings.label = TRUE, loadings.label.size = 3,
         loadings.label.repel=TRUE) 

```

autoplot(cfu.claim.pca, data=df.claims, colour="Strain.Listed", loadings = TRUE, loadings.colour = 'brown',
         loadings.label.colour='black', loadings.label = TRUE, loadings.label.size = 3,
         loadings.label.repel=TRUE) 


PCA of replicate samples  
```{r}
#nmr data
View(df.nmr)
df.pca.repnmr <- df.reps[c(6:16)]

#create pca values for subsets
repnmr.pca <- prcomp(df.pca.repnmr, scale. = TRUE)
#plot biplot with ellipses (5% CI?)
autoplot(repnmr.pca, data=df.reps, colour="Replicates", loadings = TRUE, loadings.colour = 'brown',
         loadings.label.colour='brown', loadings.label = TRUE, loadings.label.size = 4,
         loadings.label.repel=TRUE) +   ggforce::geom_mark_ellipse(aes(fill = Replicates,
                        color = Replicates))

#cfu data
df.pca.repcfu <- df.reps[c(17:20)]

#create pca values for subsets
repcfu.pca <- prcomp(df.pca.repcfu, scale. = TRUE)
#plot biplot with ellipses (5% CI?)
autoplot(repcfu.pca , data=df.reps, colour="Replicates", loadings = TRUE, loadings.colour = 'brown',
         loadings.label.colour='brown', loadings.label = TRUE, loadings.label.size = 4,
         loadings.label.repel=TRUE) +   ggforce::geom_mark_ellipse(aes(fill = Replicates,
                        color = Replicates))


```

                        
```{r}
# ttest for strain listed AND type
wilcox.test(`M13.log(CFU/ml)`~ Strain.Listed, conf.int = TRUE, exact = TRUE, data = df.claims) #W = 130, p-value = 0.1228
wilcox.test(`MRS.log(CFU/ml)`~ Strain.Listed, conf.int = TRUE, exact = TRUE, data = df.claims) #W = 74.5, p-value = 0.001796
wilcox.test(`NA.log(CFU/ml)`~ Strain.Listed, conf.int = TRUE, exact = TRUE, data = df.claims) #W = 176.5, p-value = 0.8063
wilcox.test(`WLN.log(CFU/ml)`~ Strain.Listed, conf.int = TRUE, exact = TRUE, data = df.claims) #W = 127, p-value = 0.1082
wilcox.test(Median.CFU~ Strain.Listed, conf.int = TRUE, exact = TRUE, data = df.claims) #W = 119, p-value = 0.06871

wilcox.test(`M13.log(CFU/ml)`~ Type, conf.int = TRUE, exact = TRUE, data = df.claims) #W = 138.5, p-value = 0.06322
wilcox.test(`MRS.log(CFU/ml)`~ Type, conf.int = TRUE, exact = TRUE, data = df.claims) #W = 176, p-value = 0.3752
wilcox.test(`NA.log(CFU/ml)`~ Type, conf.int = TRUE, exact = TRUE, data = df.claims) #W = 143, p-value = 0.08811
wilcox.test(`WLN.log(CFU/ml)`~ Type, conf.int = TRUE, exact = TRUE, data = df.claims) #W = 151, p-value = 0.1306
wilcox.test(Median.CFU~ Type, conf.int = TRUE, exact = TRUE, data = df.claims) #W = 119, p-value = 0.06871
#________________________________________________________

wilcox.test(`M13.log(CFU/ml)`~ Strain.Listed, conf.int = TRUE, exact = TRUE, data = df.adj) #p-value = 0.1561
wilcox.test(`MRS.log(CFU/ml)`~ Strain.Listed, conf.int = TRUE, exact = TRUE, data = df.adj) #W = 93.5, p-value = 0.0003721
wilcox.test(`NA.log(CFU/ml)`~ Strain.Listed, conf.int = TRUE, exact = TRUE, data = df.adj) #W = 176.5, p-value = p-value = 0.9182
wilcox.test(`WLN.log(CFU/ml)`~ Strain.Listed, conf.int = TRUE, exact = TRUE, data = df.adj) #W = 171, p-value = 0.07184
wilcox.test(Median.CFU~ Strain.Listed, conf.int = TRUE, exact = TRUE, data = df.adj) #W = 119, W = 165, p-value = 0.055

wilcox.test(logM13~ Type, conf.int = TRUE, exact = TRUE, data = df.adj.l) #W = 210, p-value = 0.3008
wilcox.test(logMRS~ Type, conf.int = TRUE, exact = TRUE, data = df.adj.l) #W = 166, p-value = 0.6527
wilcox.test(logNA~ Type, conf.int = TRUE, exact = TRUE, data = df.adj.l) #W = 127.5, p-value = 0.3996
wilcox.test(logWLN~ Type, conf.int = TRUE, exact = TRUE, data = df.adj.l) #W = 236, p-value = 0.1076
#wilcox.test(Median.CFU~ Type, conf.int = TRUE, exact = TRUE, data = df.adj) #W = 246, p-value = 0.1561


cor.test(df.adj$`MRS.log(CFU/ml)`, df.adj$`Acetic acid(g/L)`, method=c("spearman"))
```


Linear regression model for Median CFU and MRS CFU
```{r}
#Create new dataset with median CFU to build lineaer regression model- including AA, ethanol, claim, brand,type
#create new dataframe with variables of interest
df.linear <- df.adj[c(1,2,3,4,6,7,9,18,19,20,21)]

#determine which variables to include in regression model and look for multicollinearity
pairs(df.linear[c(6,7,8,9,10,11)])
head(df.adj)

pairs(df.adj[c(7:17,19)])
#couldn't find a model to fita linear regression- 
```
pairs(df.adj[c(7:18)])


Build factorial regression for Median CFU and MRS CFU
```{r}
#for MRS
#explore data
boxplot.type.mrs <- ggplot(df.claims.l, aes(x=Type, y=`M13.log(CFU/ml)`))
boxplot.type.mrs + geom_boxplot() #no

boxplot.claim.mrs <- ggplot(df.claims, aes(x=Claim, y=`M13.log(CFU/ml)`))
boxplot.claim.mrs + geom_boxplot() #yes
boxplot.claim.mrs <- ggplot(df.claims, aes(x=Strain.Listed,y= `M13.log(CFU/ml)`))
boxplot.claim.mrs + geom_boxplot() #yes

df.claims$Claim <- as.factor(df.claims$Claim)
MRSModel<-aov(`MRS.log(CFU/ml)` ~ Strain.Listed * Claim, data = df.claims)
summary(MRSModel)

plot(MRSModel, 1)
leveneTest(`MRS.log(CFU/ml)` ~ Strain.Listed * Claim, data = df.claims)

plot(MRSModel, 2)
#for Median CFU
TukeyHSD(MRSModel, which = "Strain.Listed")

library(multcomp)
summary(glht(MRSModel, linfct = mcp(Claim = "Tukey")))

MedianModel<-aov(Median.CFU ~ Claim * Strain.Listed, data = df.claims)
summary(MedianModel)
```


```{r}
# create the lollipop plot

df.readcounts <-  read_csv("Lac.Bac.Counts.csv")
df.readcounts <- df.readcounts[,-1]
df.bacillus <- df.readcounts %>% filter(Organism == "Bacillus")
df.lactobacillus <- df.readcounts %>% filter(Organism == "Lactobacillus")


# DATAFRAME df.readcounts dataframe with read counts and categories
df.readcounts[is.na(df.readcounts)] <- 0 

df.readcounts$Reads <- factor(df.readcounts$Reads, levels=c("High", "Med", "Low"))

p <- ggplot(df.readcounts, aes(y = Reads , x = `logNA`, fill = Organism, color=Organism)) + geom_segment(aes(x = 0, y = Reads, xend = `logNA`, yend = Reads), color = "grey") + geom_point(color = "orange") + facet_wrap(~Organism) +theme_bw()

p2 <- ggplot(df.readcounts, aes(y = Reads , x = Max.cfu, fill = Organism, color=Organism)) + geom_segment(aes(x = 0, y = Reads, xend = Max.cfu, yend = Reads), color = "grey") + geom_point(color = "blue") + facet_wrap(~Organism) +theme_bw()

p3 <- ggplot(df.readcounts, aes(y = Reads , x = `logMRS`, fill = Organism, color=Organism)) + geom_segment(aes(x = 0, y = Reads, xend = `logMRS`, yend = Reads), color = "grey") + geom_point(color = "red") + facet_wrap(~Organism) +theme_bw()

#--------Perform ANOVA for stats
kruskal.test(`logMRS` ~ Reads, data = df.bacillus) #Kruskal-Wallis p-value = 0.06258
kruskal.test(`logNA` ~ Reads, data = df.bacillus) #Kruskal-Wallis p-value = 0.173
kruskal.test(Max.cfu ~ Reads, data = df.bacillus) #Kruskal-Wallis p-value = 0.1987

kruskal.test(`logMRS` ~ Reads, data = df.lactobacillus) #Kruskal-Wallis p-value = 0.01385
dunnTest(`logMRS` ~ Reads, data = df.lactobacillus, method="bonferroni") #**0.01299504 low-med
kruskal.test(`logNA` ~ Reads, data = df.lactobacillus) #Kruskal-Wallis p-value = 0.07757
kruskal.test(Max.cfu ~ Reads, data = df.lactobacillus) #Kruskal-Wallis p-value = 0.001671
dunnTest(Max.cfu ~ Reads, data = df.lactobacillus, method="bonferroni") #0.006649067 hi-low 0.008666620 #low-medium 
```


Looking at consistencies within replicates 
```{r}
#df with just the replicated samples
df.reps<- df.snpro %>% drop_na("Replicates") #remove samples that do not have a replicate
df.reps[is.na(df.reps)] <- 0 #convert NA to 0
df.reps.2 <- df.reps[c(4,11:21,31,29,27,25,23)]

#create a violin plot of CFU distribution for replicates
M13.r <- ggplot(df.reps.2, aes(x= Replicates, y=`logM13`, fill=Replicates)) +
  geom_violin()+
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") + scale_x_discrete(guide = guide_axis(check.overlap = TRUE))
MRS.r <- ggplot(df.reps.2, aes(x= Replicates, y=`logMRS`, fill=Replicates)) +
  geom_violin()+
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") + scale_x_discrete(guide = guide_axis(check.overlap = TRUE))
WLN.r <- ggplot(df.reps.2, aes(x= Replicates, y=`logWLN`, fill=Replicates)) +
  geom_violin()+
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") + scale_x_discrete(guide = guide_axis(check.overlap = TRUE))
NAm.r <- ggplot(df.reps.2, aes(x= Replicates, y=`logNA`, fill=Replicates)) +
  geom_violin()+
  geom_boxplot(width=0.1) + scale_fill_brewer(palette="RdPu") + scale_x_discrete(guide = guide_axis(check.overlap = TRUE))

#---------------Make biplot
df.reps.NMR <- df.reps[c(11:21)]
#create pca values for subsets
reps.pca <- prcomp(df.reps.NMR, scale. = TRUE)
#plot biplot with ellipses (5% CI?)
autoplot(reps.pca , data=df.reps, colour="Replicates", loadings = TRUE, loadings.colour = 'black',
         loadings.label.colour='brown', loadings.label = TRUE, loadings.label.size = 3,
         loadings.label.repel=TRUE) +   ggforce::geom_mark_ellipse(aes(fill = Replicates,
                        color = Replicates))
```

autoplot(reps.pca , data=df.reps, colour="Replicates", loadings = TRUE, loadings.colour = 'brown',
         loadings.label.colour='black', loadings.label = TRUE, loadings.label.size = 3,
         loadings.label.repel=TRUE) +   ggforce::geom_mark_ellipse(aes(fill = Replicates,
                        color = Replicates)) + theme_bw()


```{r}
bottle.df <- read_csv("Bottle.Diff.Bar.csv") #dataframe containing comparison of CFU count on bottle vs observed SPC
head(bottle.df)
bottle.df

bottle.df$Value <- factor(bottle.df$Value, levels=c("Max:SPC", "Difference"))

#Default bar plot
#fill= Brand; x-axis= Value, y-axis= 'log(CFU ml-1)'
p<- ggplot(bottle.df, aes(x=Brand, y=`log(CFU ml-1)`, fill=Value)) + 
  geom_bar(stat="identity", color="black", 
           position=position_dodge()) +
  geom_errorbar(aes(ymin=`log(CFU ml-1)`-SD, ymax=`log(CFU ml-1)`+SD), width=.2,
                 position=position_dodge(.9)) + theme_bw() +
scale_fill_brewer(palette="Paired") + scale_color_brewer(palette="Paired") + theme(legend.position = "bottom")

print(p)
```
print(p)

Arrange images for figure 2 + 3 
```{r}
#figure2
ggarrange(p6, p7,nrow = 1,ncol=2, common.legend = TRUE, legend = "bottom")
#figure3
ggarrange(hist.cfu.type, hist.cfu.claim,organism,p,common.legend = FALSE, legend = "bottom")

hist.cfu.type/(WLN.t|M13.t)/p
```

ggarrange(p6, p7,nrow = 1,ncol=2, common.legend = TRUE, legend = "bottom")
ggarrange(hist.cfu.type, hist.cfu.claim,organism,p,common.legend = FALSE, legend = "bottom")

MANOVA comparing NMR of types 
```{r}
#independent variable- type
#dependent variables- compounds without ethanol 

#create dataset with just type and compounds of interest #method from https://www.r-bloggers.com/2022/01/manova-in-r-how-to-implement-and-interpret-one-way-manova/

nmr.MANOVA <- df.adj[c(2,10,11,14:20)]
head(nmr.MANOVA)
#combine dependent variables: Acetic acid(g L-1), Erythritol(g L-1), Fructose(g L-1) Gluconic acid(g L-1), Glucose(g L-1), Isobutyrate(mg L-1), Lactic acid(g L-1), Sucrose(g L-1)
nmr.MANOVA <- na.omit(nmr.MANOVA)

dependent_vars <- cbind(nmr.MANOVA$`Acetic acid(g L-1)`, nmr.MANOVA$`Erythritol(g L-1)`, nmr.MANOVA$`Fructose(g L-1)`, nmr.MANOVA$`Gluconic acid(g L-1)`, nmr.MANOVA$`Glucose(g L-1)`, nmr.MANOVA$`Glycerol(g L-1)`, nmr.MANOVA$`Isobutyrate(mg L-1)`, nmr.MANOVA$`Lactic acid(g L-1)`, nmr.MANOVA$`Sucrose(g L-1)`)

#set independent variable
independent_var <- nmr.MANOVA$Type

#create manova model 
manova_model <- manova(dependent_vars ~ independent_var, data = nmr.MANOVA)
summary(manova_model)

#Post-hoc test by performing a linerar discriminant analysis (from: https://www.reneshbedre.com/blog/manova.html)

post_hoc <- lda(independent_var ~ dependent_vars, CV=F)
post_hoc

#plot
plot_lda <- data.frame(nmr.MANOVA[, "Type"], lda = predict(post_hoc)$x)

MANOVA.plot <- ggplot(plot_lda) + geom_point(aes(x = lda.LD1, y = lda.LD2, colour = Type), size = 2) + theme_bw() + stat_ellipse(geom = "polygon",
               aes(x = lda.LD1, y = lda.LD2, fill = Type), 
               alpha = 0.25)

print(MANOVA.plot)
```
print(MANOVA.plot)

Run another MANOVA comparing soft to soft-aberrant 
```{r}
#independent variable- type
#dependent variables- compounds without ethanol 
SNProR_soft <- read_csv("SNProR.soft.csv", 
    na = "empty")
#remove SN#ID and ethanol concentration
Soft.MANOVA.nmr <- SNProR_soft[-c(1,4)]

#create dataset with just type and compounds of interest #method from https://www.r-bloggers.com/2022/01/manova-in-r-how-to-implement-and-interpret-one-way-manova/

#combine dependent variables: Acetic acid(g L-1) Fructose(g L-1) Gluconic acid(g L-1), Glucose(g L-1), Isobutyrate(mg L-1), Lactic acid(g L-1), Sucrose(g L-1)

dependent_vars <- cbind(Soft.MANOVA.nmr$`Acetic acid(g L-1)`,Soft.MANOVA.nmr$`Fructose(g L-1)`, Soft.MANOVA.nmr$`Gluconic acid(g L-1)`, Soft.MANOVA.nmr$`Glucose(g L-1)`, Soft.MANOVA.nmr$`Glycerol(g L-1)`, Soft.MANOVA.nmr$`Isobutyrate(mg L-1)`, Soft.MANOVA.nmr$`Lactic acid(g L-1)`, Soft.MANOVA.nmr$`Sucrose(g L-1)`)

#set independent variable
independent_var <- Soft.MANOVA.nmr$Type

#create manova model 
manova_model_2 <- manova(dependent_vars ~ independent_var, data = Soft.MANOVA.nmr)
summary(manova_model_2)


```


correlation plot
```{r}

##function for p-values
cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}


##make df for corrmatrix
df.corr <- df.adj[10:22]
head(df.corr)
df.corr <- df.corr[,-4]
df.corr[is.na(df.corr)] <- 0.0000001

corr.matrix <-cor(df.corr,use="complete.obs", method="pearson") #calculate correlation values 
p.mat <- cor.mtest(corr.matrix)
head(p.mat)
corr <- corrplot(corr.matrix, method="circle", type="upper", order="hclust", col=brewer.pal(n=8, name="RdYlBu"), 
addCoef.col = "black",
         tl.col="black", tl.srt=45, number.cex = 0.5,
p.mat = p.mat, sig.level = 0.01, insig = "blank")



# Create the correlation plot
corr <- corrplot(corr.matrix, method="circle", type="upper", order="hclust", col=brewer.pal(n=8, name="RdYlBu"), 
addCoef.col = "black",
         tl.col="black", tl.srt=45)

# Add the p-values to the plot
for (i in 1:ncol(p.mat)) {
    for (j in 1:nrow(p.mat)) {
        if (i > j) {
            text(x = i, y = j, labels = format(p.mat[i, j], scientific = TRUE), number.cex = 0.2, cex.axis = 0.1, font = 0.2, xpd = TRUE)
        }
    }
}

```
corr

```{r}

df.replicates <- read_csv("snopro.reps.csv") #original dataset containing demograohic, cfu, logcfu, and NMR data- no samples removed

df.rep.nmr <- df.replicates[c(3,6:18)]
df.rep.nmr[is.na(df.rep.nmr)] <- 0.00001

#create pca values for subsets
nmr.pca <- prcomp(df.rep.nmr[c(2:13)], scale. = TRUE)
#plot biplot with ellipses (5% CI?)
autoplot(nmr.pca, data=df.rep.nmr, colour="Replicate", loadings = TRUE, loadings.colour = 'brown',
         loadings.label.colour='brown', loadings.label = TRUE, loadings.label.size = 3,
         loadings.label.repel=TRUE) 


library(devtools)
install_github("vqv/ggbiplot")
# Perform PCA on your data
pca_results <- prcomp(df.rep.nmr[c(2:13)], scale = TRUE)

# Create a ggbiplot object
pca_biplot <- ggbiplot(pca_results, obs.scale = 1, var.scale = 1, groups = df.rep.nmr$Replicate,ellipse = TRUE, ellipse.level = 0.95)

# Customize the plot
pca_biplot + theme_classic() +  scale_color_discrete(name = "Replicate")

library(factoextra)
fviz_pca_ind(pca_results, 
             geom = "point", 
             ellipse.type = "confidence", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             addEllipses = TRUE,
             label = "var",
             repel = TRUE)

```

